---
title: "Data Preparation for Case 31"
subtitle: "Complete Version"
author: "Jae Jung"
institute: "IBM 3302, Cal Poly Pomona"
date: "`r Sys.Date()`"
format: 
  revealjs:
    footer: Case 31 Project
    transition: slide
    transition-speed: default
    toc: true
    toc-depth: 1
    slide-number: true
    scrollable: true
    smaller: true
    code-fold: false
    code-overflow: wrap
    code-copy: true
    number-sections: true
    embed-resources: true
    self-contained-math: true
editor: visual
execute: 
  echo: true
  freeze: auto
---

# Libraries {background-color="aquamarine"}

## Loading common packages

```{r}
# install.packages("haven")
library(tidyverse)
library(haven) 
library(labelled) 
library(sjlabelled)
library(janitor) 
library(sjPlot)
library(psych)
set_theme(theme_minimal())
```

# Setting Up Data {background-color="aquamarine"}

## Reading SPSS data {.smaller}

```{r}
#| output-location: fragment

data <- read_sav("Case 31 Data_efcu_1.sav")
data <- data |> 
  clean_names()

# generate data dictionary
data |>
  view_df()

```

# Data Preparation {background-color="aquamarine"}

## Attitudes towards Employee Proficiency {.smaller}

### q3

```{r}
data |> # ctr + shift + M
  count(q3) |> 
  mutate(percent = n/sum(n))
```

------------------------------------------------------------------------

```{r}
data_clean <- data |> 
  set_value_labels(q3 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5)) |> 
  set_variable_labels(q3 = "Employees are Courteous")

data_clean |> 
  select(q3) |> 
  var_label() 

data_clean |> 
  select(q3) |> 
  val_labels()
```

------------------------------------------------------------------------

### q4 {.smaller}

```{r}
data_clean <-  data_clean |> 
  set_variable_labels(q4 = "Employees are helpful") |> 
  set_value_labels(q4 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5)) 

data_clean |> 
  select(q4) |>
  var_label()

data_clean |> 
  select(q4) |> 
  val_labels()
```

------------------------------------------------------------------------

### q5 and q6

```{r}
data_clean <- data_clean |> 
  set_variable_labels(q5 = "Employees are professional") |> 
  set_value_labels(q5 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5)) |> 
  set_variable_labels(q6 = "Employees are available") |> 
  set_value_labels(q6 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5))
data_clean |> 
  select(q5, q6) |>
  var_label()

data_clean |> 
  select(q5, q6) |> 
  val_labels()
```

------------------------------------------------------------------------

### Reliability analysis

```{r}
#| output-location: fragment

data_clean |> 
  select(q3:q6) |> 
  alpha() #.87: cutoff is .70
```

------------------------------------------------------------------------

### factor analysis

```{r}
data_clean |> 
  select(q3:q6) |> 
  factanal(factors = 1, rotation = "none")

data_clean |> 
  select(q3:q6) |> 
  fa(nfactors = 1, rotate = "none") |> 
  fa.diagram()

```

------------------------------------------------------------------------

### PCA (Principle Component Analysis)

```{r}
data_clean |>
  select(q3:q6) |> 
  prcomp(scale. = TRUE) |> 
  summary() # % of variance explained by the first factor
```

------------------------------------------------------------------------

### Scree plot

```{r}
data_clean |> 
  select(q3:q6) |> 
  scree()
```

------------------------------------------------------------------------

### Creating a composite scale

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(at_emp = rowMeans(across(q3:q6), na.rm = TRUE)) 

# from psych: descriptive statistics
describe(data_clean$at_emp)

```

## q7r {.smaller}

::::: columns
::: {.column width="40%"}
```{r}
data_clean |> count(q7)
```
:::

::: {.column width="60%"}
```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q7r = case_when(q7 == 1 ~ 5,
                         q7 == 2 ~ 4,
                         q7 == 3 ~ 3,
                         q7 == 4 ~ 2,
                         q7 == 5 ~ 1,
                         TRUE    ~ NA)) |> 
  set_variable_labels(q7r = "Savings rates") |> 
  set_value_labels(q7r = c("Very Low" = 1,
                           "Low" = 2,
                           "Average" = 3,
                           "High" = 4,
                           "Very High" = 5))

data_clean |> count(q7r)
```
:::
:::::

## q8r

```{r}
#| output-location: fragment

data_clean |> count(q8)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q8r = case_when(q8 == 1 ~ 5,
                         q8 == 2 ~ 4,
                         q8 == 3 ~ 3,
                         q8 == 4 ~ 2,
                         q8 == 5 ~ 1,
                         TRUE    ~ NA)) |> 
  set_variable_labels(q8r = "Fund borrowing rates") |> 
  set_value_labels(q8r = c("Very Low" = 1,
                           "Low" = 2,
                           "Average" = 3,
                           "High" = 4,
                           "Very High" = 5))

data_clean |> count(q8r)
```

------------------------------------------------------------------------

## q9r

```{r}
#| output-location: fragment

data_clean |> count(q9)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean |> 
  select(q9) |> 
  class()
data_clean <- data_clean |> 
  mutate(q9r = 6 - q9) |> 
  set_variable_labels(q9r = "Statement Frequency") |> 
  set_value_labels(q9r = c("Never" = 1,
                           "Not Often Enough" = 2,
                           "About Right" = 3,
                           "Very Often" = 4,
                           "Too Often" = 5))
data_clean |> count(q9r)  
```

------------------------------------------------------------------------

## q10r

```{r}
#| output-location: fragment

data_clean |> count(q10)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q10r = 5 - q10) |> 
  set_variable_labels(q10r = "Statement Accuracy") |> 
  set_value_labels(q10r = c("Poor" = 1,
                           "Fair" = 2,
                           "Good" = 3,
                           "Excellent" = 4))
data_clean |> count(q10r)  
```

## q13r

```{r}
#| output-location: fragment

data_clean |> count(q13)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q13r = 4 - q13) |> 
  set_variable_labels(q13r = "Regular share account") |> 
  set_value_labels(q13r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q13r)  
```

## q14r

```{r}
#| output-location: fragment

data_clean |> count(q14)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q14r = 4 - q14) |> 
  set_variable_labels(q14r = "Special subaccounts") |> 
  set_value_labels(q14r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q14r)  

```

## q15r

```{r}
#| output-location: fragment

data_clean |> count(q15)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q15r = 4 - q15) |> 
  set_variable_labels(q15r = "Christmas club account") |> 
  set_value_labels(q15r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q15r)  
```

## q16r

```{r}
#| output-location: fragment

data_clean |> count(q16)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q16r = 4 - q16) |> 
  set_variable_labels(q16r = "IRA") |> 
  set_value_labels(q16r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q16r)  
```

## q17r

```{r}
#| output-location: fragment

data_clean |> count(q17)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q17r = 4 - q17) |> 
  set_variable_labels(q17r = "Master credit card") |> 
  set_value_labels(q17r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q17r)  
```

## q18r

```{r}
#| output-location: fragment

data_clean |> count(q18)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q18r = 4 - q18) |> 
  set_variable_labels(q18r = "Signature loans") |> 
  set_value_labels(q18r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q18r)  
```

## q19r

```{r}
#| output-location: fragment

data_clean |> count(q19)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q19r = 4 - q19) |> 
  set_variable_labels(q19r = "New car loans") |> 
  set_value_labels(q19r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q19r)  
```

## q20r

```{r}
#| output-location: fragment

data_clean |> count(q20)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q20r = 4 - q20) |> 
  set_variable_labels(q20r = "Late model car loans") |> 
  set_value_labels(q20r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q20r)  
```

## q21r

```{r}
#| output-location: fragment

data_clean |> count(q21)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q21r = 4 - q21) |> 
  set_variable_labels(q21r = "Older model car loans") |> 
  set_value_labels(q21r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q21r)  
```

## q22r

```{r}
data_clean |> count(q22)

data_clean <- data_clean |> 
  mutate(q22r = 4 - q22) |> 
  set_variable_labels(q22r = "Household goods loans") |> 
  set_value_labels(q22r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q22r)  
```

## q23r

```{r}
data_clean |> count(q23)

data_clean <- data_clean |> 
  mutate(q23r = 4 - q23) |> 
  set_variable_labels(q23r = "Recreational loans") |> 
  set_value_labels(q23r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q23r)  
```

## q24r

```{r}
data_clean |> count(q24)

data_clean <- data_clean |> 
  mutate(q24r = 4 - q24) |> 
  set_variable_labels(q24r = "Share collateralized loan") |> 
  set_value_labels(q24r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q24r)  
```

## q25r

```{r}
data_clean |> count(q25)

data_clean <- data_clean |> 
  mutate(q25r = 4 - q25) |> 
  set_variable_labels(q25r = "IRA loans") |> 
  set_value_labels(q25r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q25r)  
```

## q26r

```{r}
data_clean |> count(q26)

data_clean <- data_clean |> 
  mutate(q26r = 4 - q26) |> 
  set_variable_labels(q26r = "Line of credit loans") |> 
  set_value_labels(q26r = c("Unaware" = 1,
                           "Aware but Have Not Used" = 2,
                           "Aware and Have Used" = 3))
data_clean |> count(q26r)  
```

## q30 - q34 {.smaller}

```{r}
data_clean <-  data_clean |> 
  set_value_labels(q30 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5)) |> 
  set_value_labels(q31 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5))|> 
  set_value_labels(q32 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5))|> 
  set_value_labels(q33 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5))|> 
  set_value_labels(q34 = c("Strongly Disagree" = 1,
                          "Disagree" = 2,
                          "Uncertain" = 3,
                          "Agree" = 4,
                          "Strongly Agree" = 5))
```

## q37r

```{r}
#| output-location: fragment

data_clean |> count(q37)
```

------------------------------------------------------------------------

```{r}
#| output-location: fragment

data_clean <- data_clean |> 
  mutate(q37r = case_when(q37 == 1 ~ 5,
                          q37 == 2 ~ 4,
                          q37 == 3 ~ 3,
                          q37 == 4 ~ 2,
                          q37 == 5 ~ 1,
                          q37 == 6 ~ NA)) |> 
  set_value_labels(q37r = c("Very Poor" = 1,
                            "Poor" = 2,
                            "Average" = 3,
                            "Good" = 4,
                            "Excellent" = 5))

data_clean |> count(q37r)
```

# Renaming Clean Data

```{r}
data_preped <- data_clean

data_preped |> 
  view_df()
```

# Exporting Prepped Data

## SPSS data

```{r}
#| eval: false

data_preped |> 
  write_sav("Case 31 Data_efcu_Prepped.sav")

```

## Excel Data as Metadata

```{r}
#| eval: false

library(writexl)

data_preped |> 
  write_xlsx("Case 31 Data_efcu_Prepped.xlsx")

```

```{r}
#| eval: false

# Extract variable labels
var_labels <- get_label(data)
var_labels <- var_labels |> 
  t() |> 
  as.data.frame()
dim(var_labels)
var_labels

# Convert data to labeled values
labeled_data <- as_label(data)
dim(labeled_data)
labeled_data <- labeled_data |> 
  as.data.frame()

# Add labels as the first row
output_data <- bind_rows(
  var_labels, labeled_data)

# Save to Excel
write_xlsx(output_data, path = "Case 31 Data_efcu_Prepped_labels.xlsx")

```

# Appendix

## alternative way using select(): in this case, we need to use anonymous function

```{r}
#| eval: false
#| echo: true

data_clean |>
  (\(df) df |> 
    mutate(row_mean = rowMeans(select(df, q3, q4, q5, q6), na.rm = TRUE))
  )() |> 
  select(row_mean)
```
